---
title: "NEW_Growth Metrics_Models 041421"
output:
  html_document:
    df_print: paged
---


SVL Data: SVL's on femoralis tads (from both single- and mixed species assemblages)

```{r}
#loading packages
library(predictionInterval)
library(coefplot2)
library(merTools)
#library(AICcmodavg)
library(brms)
library(utils)
library(dbplyr)
library(TMB)
library(glmmTMB)
library(stats)
library(finalfit)
library(boot)
library(lme4)
library(MCMCglmm)
library(blme)
library(pbkrtest)
library(coda)
library(aods3)
library(bbmle)
library(ggplot2)
library(reshape2)
library(plyr)
library(numDeriv)
library(Hmisc)
library(plotMCMC)
library(gridExtra)
library(R2admb)
library(tibble)
library(tidyverse)
library(car)
library(MASS)
library(stats4)
library(readr)


SVL <- read.csv(file = "~/Desktop/SVL_Fem.csv",
                             header = TRUE,
                             sep = ",",
                             stringsAsFactors = FALSE)
```


```{r}
theme_set(theme_bw())
library("grid")          ## for unit() (base R)
zmargin <- theme(panel.margin=unit(0,"lines")) ## to squash facets together ...
```

```{r}
library(blme)
library(car)
library(nnet)
library(MASS)


#running our model
m1_svl2 <- bglmer(svl_cm~assem_type*ind_type+(1|tank_num),data=SVL,family=gaussian(link=log))
m1b_svl2 <- blmer(log(svl_cm)~assem_type*ind_type+(1|tank_num),data=SVL)
ICtab(m1_svl2,m1b_svl2,type="AICc")
summary(m1b_svl2)
Anova(m1b_svl2, type="III")
```

```{r}
#Diagnostics
plot(m1b_svl2)
qqnorm(resid(m1b_svl2))
qqline(resid(m1b_svl2))
library(visreg)
visreg(m1_svl2, "ind_type" ,by = "assem_type")
```

```{r}

SVL$ind_type=as.factor(SVL$ind_type)
SVL$assem_type=as.factor(SVL$assem_type)
SVL$trmt_ID=factor(SVL$trmt_ID, levels=c("NL_C_S", "C_C_S", "NL_C_M", "C_C_M"))
str(SVL)


library(dotwhisker)
dotwhisker::dwplot(m1_svl2,effects="fixed",by_2sd=FALSE)
#rr <- "http://www.math.mcmaster.ca/bolker/R"
#install.packages(c("coefplot2"),type="source",
                # repos=rr)
coeftab(m1b_svl2)[,c("Estimate","2.5%","97.5%")]
coefplot2(m1b_svl2)
predict(m1b_svl2,type="response",se=TRUE)
```

```{r}
dataset=SVL
boot_predict <- function(.) {
  predict(., newdata=dataset, re.form=NULL)
}



#install.packages('TMB', type = 'source')
bootSummary <- function(merBoot) {
  return(
    data.frame(fit = apply(merBoot$t, 2, function(x) as.numeric(quantile(x, probs=.5, na.rm=TRUE))),
               lwr = apply(merBoot$t, 2, function(x) as.numeric(quantile(x, probs=.025, na.rm=TRUE))),
               upr = apply(merBoot$t, 2, function(x) as.numeric(quantile(x, probs=.975, na.rm=TRUE)))
    )
  )
}

merboot=bootMer(m1b_svl2, boot_predict, nsim=200, use.u=FALSE, type="parametric")
bootstrap.PIs <- bootSummary(merboot)
pframe=data.frame(dataset,bootstrap.PIs)
library(dplyr)
pframe %>% 
  group_by(ind_type, assem_type) %>% 
  summarise(fit = mean(fit),lwr=mean(lwr),upr=mean(upr)) -> pframe2


#renaming factors for plots
levels(pframe2$assem_type) <-list(Mixed="M",Single="S")
levels(pframe2$assem_type)
levels(pframe2$ind_type) <-list(Nonlethal="NL", Control="C")
levels(pframe2$ind_type)

#transforming logged data
pframe2$expfit <- exp(pframe2$fit)
pframe2$explwr <- exp(pframe2$lwr)
pframe2$expupr <- exp(pframe2$upr)
pframe2

```


```{r}
library(ggplot2)
ggplot(aes(x=ind_type, y=expfit, ymin=explwr, ymax=expupr, color=assem_type), data=pframe2) +
  geom_point(size=4,position = position_dodge(w = 0.75))+
  geom_errorbar(position = position_dodge(w = 0.75)) +
  geom_line(aes(group=assem_type),position = position_dodge(w = 0.75))+
  ylim(1.00,1.40)+
  labs(colour="Assemblage Type",x="Induction Treatment", y="SVL (cm) w/ 95% PIs") +
  theme_bw() +  theme(legend.position="right") +
  scale_color_brewer(type = "qual", palette = 2)
```

Plotting the raw SVL data:

```{r}
library(ggplot2)
library(scales)

ind_labs <- c("Control", "Non-Lethal")
names(ind_labs) <- c("C", "NL")


ggplot(SVL, aes(x=assem_type, y=svl_cm, fill=assem_type))+
  geom_boxplot(size=.75, point=TRUE)+
  geom_jitter(aes(colour=assem_type), alpha=0.3)+
  scale_fill_manual(values = c("darkseagreen4","darkorange2"))+
  stat_summary(fun.y = "mean", geom = "point", shape = 15, size = 2)+
  labs(title="Femoralis Snout-Vent Lengths", x="Assemblage Type", y="Femoralis SVL (cm)")+
  facet_wrap(~ind_type, labeller = labeller(ind_type = ind_labs))+
  scale_color_brewer(type="qual", palette=2)+
  theme_bw()
```


MASS Data: Emergence-day mass -


```{r}
Emass <- read.csv(file = "~/Desktop/Masses_Fem.csv",
                             header = TRUE,
                             sep = ",",
                             stringsAsFactors = FALSE)


#Run the model
m1_mass <- glmer(emerg_grams~assem_type*ind_type+(1|tank_num),data=Emass,family=gaussian(link=log))
summary(m1_mass)
Anova(m1_mass, type="III")
```

```{r}
#plotting residuals & model diagnostics - Looks acceptable to me! Generally homoschedastic
plot(m1_mass,idLabels=~.obs,id=.1)
qqnorm(resid(m1_mass))
qqline(resid(m1_mass))

#plotting residuals without outlier #9


library(visreg)
visreg(m1_mass, "ind_type" ,by = "assem_type")
##Observation 9 looks to be an outlier
Emass[7:10,]
```

```{r}
#Run the model with outlier removed - need to check on it: Outlier=9
new_Emass <- data.frame(Emass[-9,])
m1b_mass <- lmer(log(emerg_grams)~assem_type*ind_type+(1|tank_num),data=new_Emass)
summary(m1b_mass)
Anova(m1b_mass)
m1c_mass <- glmer((emerg_grams)~assem_type*ind_type+(1|tank_num),data=new_Emass,family=gaussian(link=log))
ICtab(m1b_mass,m1c_mass,type="AICc")
```


```{r}
#plotting residuals without outlier #9
# I know have outliers 34, 35 and 47 as well 
plot(m1b_mass,idLabels=~.obs,id=.1)
qqnorm(resid(m1b_mass))
qqline(resid(m1b_mass))
```
https://www.ssc.wisc.edu/sscc/pubs/MM/MM_DiagInfer.html

https://www.ssc.wisc.edu/sscc/pubs/RFR/RFR_Diagnostics.html

https://dfzljdn9uc3pi.cloudfront.net/2020/9522/1/MixedModelDiagnostics.html

```{r}
#diagnostics for leverage
#need to write a model WITH the outlier in it first 
new_mass <- lmer(log(emerg_grams)~assem_type*ind_type+(1|tank_num),data=Emass)
summary(new_mass)

#viewing influence plots of new_mass that includes ALL observations
library(influence.ME)
infl <- influence(new_mass, obs=TRUE)
plot(infl)
plot(infl,which="cook")
influencePlot(new_mass)

#viewing influence plots with outlier #9 removed
infl_9 <- influence(m1b_mass,obs=TRUE)
plot(infl_9,which="cook")
influencePlot(m1b_mass)



#new model with additional outliers #34, 35 & 47 removed
newEmass2 <- data.frame(new_Emass[-34,])
newEmass3 <- data.frame(newEmass2[-33,])
newEmass4 <- data.frame(newEmass3[-44,])
new_mass2 <- lmer(log(emerg_grams)~assem_type*ind_type+(1|tank_num),data=newEmass4)
summary(new_mass2)
Anova(new_mass2)

#plotting influence with outliers #34, 35 & 47 removed
#plots look like I'm chasing the outlier bubble....
library(influence.ME)
infl_outs <- influence(new_mass2,obs=TRUE)
plot(infl_outs,which="cook")

```


```{r}
##although the AIC is better for the glmer, diagnostics below suggest that the lmer on log-transformed data is better due to nonlinearity rather than just log normal error structure
summary(m1c_mass)
fixef(m1b_mass)
fixef(m1c_mass)
Anova(m1c_mass, type="III")
Anova(m1b_mass, type="III")
```


```{r}
par(mfrow=c(1,2))
plot(m1c_mass,idLabels=~.obs,id=.1)
plot(m1b_mass,idLabels=~.obs,id=.1,title="log(mass)")
par(mfrow=c(1,2))
qqnorm(resid(m1c_mass))
qqline(resid(m1c_mass))
qqnorm(resid(m1b_mass),title="log(mass)")
qqline(resid(m1b_mass))
```


```{r}
sc1=ggplot(fortify.merMod(m1b_mass),
       aes(x=.fitted,y=sqrt(abs(.scresid))))+geom_point()+
    geom_smooth(colour="red",alpha=0.3)+ggtitle("log(mass")
sc2=ggplot(fortify.merMod(m1c_mass),
       aes(x=.fitted,y=sqrt(abs(.scresid))))+geom_point()+
    geom_smooth(colour="red",alpha=0.3)
grid.arrange(sc1,sc2,ncol=2)
library(visreg)
visreg(m1b_mass, "ind_type" ,by = "assem_type")
```

```{r}
Emass$ind_type=as.factor(Emass$ind_type)
Emass$assem_type=as.factor(Emass$assem_type)
Emass$trmt_ID=factor(Emass$trmt_ID, levels=c("NL_C_S", "C_C_S", "NL_C_M", "C_C_M"))
str(Emass)


#a look at Regression estimates, with and without random effects. Nothing significant here
library(dotwhisker)
library(coefplot2)
dotwhisker::dwplot(m1b_mass,effects="fixed",by_2sd=FALSE)
#rr <- "http://www.math.mcmaster.ca/bolker/R"
#install.packages(c("coefplot2"),type="source",
                # repos=rr)
coeftab(m1b_mass)[,c("Estimate","2.5%","97.5%")]
coefplot2(m1_mass)
predict(m1b_mass,type="response",se=TRUE)
```

```{r}
#predicted values and confidence intervals
dataset1=Emass
boot_predict <- function(.) {
  predict(., newdata=dataset1, re.form=NULL)
}


library(TMB)
bootSummary <- function(merBoot) {
  return(
    data.frame(fit = apply(merBoot$t, 2, function(x) as.numeric(quantile(x, probs=.5, na.rm=TRUE))),
               lwr = apply(merBoot$t, 2, function(x) as.numeric(quantile(x, probs=.025, na.rm=TRUE))),
               upr = apply(merBoot$t, 2, function(x) as.numeric(quantile(x, probs=.975, na.rm=TRUE)))
    )
  )
}



merboot=bootMer(m1b_mass, boot_predict, nsim=200, use.u=FALSE, type="parametric")
bootstrap.PIs <- bootSummary(merboot)
mass.frame=data.frame(dataset1,bootstrap.PIs)
library(dplyr)
mass.frame %>% 
  group_by(ind_type, assem_type) %>% 
  summarise(fit = mean(fit),lwr=mean(lwr),upr=mean(upr)) -> mass.frame2

#renaming factors for plots
levels(mass.frame2$assem_type)<-list(Mixed="M",Single="S")
print(mass.frame2)
levels(mass.frame2$ind_type)<-list(Nonlethal="NL", Control="C")
mass.frame2

#transforming logged data
mass.frame2$expfit <- exp(mass.frame2$fit)
mass.frame2$explwr <- exp(mass.frame2$lwr)
mass.frame2$expupr <- exp(mass.frame2$upr)
mass.frame2
```


```{r}
library(ggplot2)
ggplot(aes(x=ind_type, y=expfit, ymin=explwr, ymax=expupr, color=assem_type), data=mass.frame2) +
  geom_point(size=4,position = position_dodge(w = 0.75))+
  geom_errorbar(position = position_dodge(w = 0.75)) +
  geom_line(aes(group=assem_type),position = position_dodge(w = 0.75))+
  ylim(.15,.40)+
  labs(colour="Assemblage Type",x="Induction Treatment", y="Mass (g) w/ 95% PIs") +
  theme_bw() +  theme(legend.position="right") +
  scale_color_brewer(type = "qual", palette = 2)

```
Plotting the raw Emergence-Date mass data: 

```{r}
library(ggplot2)
library(scales)

ind_labs <- c("Control", "Non-Lethal")
names(ind_labs) <- c("C", "NL")


ggplot(Emass[-9,], aes(x=assem_type, y=emerg_grams, fill=assem_type))+
  geom_boxplot(size=.75, point=TRUE)+
  geom_jitter(aes(colour=assem_type), alpha=0.3)+
  scale_fill_manual(values = c("darkseagreen4","darkorange2"))+
  stat_summary(fun.y = "mean", geom = "point", shape = 15, size = 2)+
  labs(title="Femoralis Emergence-Date Masses", x="Assemblage Type", y="Mass (g)")+
  facet_wrap(~ind_type, labeller = labeller(ind_type = ind_labs))+
  scale_color_brewer(type="qual", palette=2)+
  theme_bw()
```


TIME-TO Emergence Data:


```{r}
emerge <- read.csv(file = "~/Desktop/Emerge Dates_Fem.csv",
                             header = TRUE,
                             sep = ",",
                             stringsAsFactors = FALSE)

view(emerge)
head(emerge)

#checking to see how many zero's we're dealing with 
df <- emerge %>%
  group_by(cumul_count) %>%
  summarise(counts = n())
df
```
A little high in the number of 0's...
```{r}
#model and diagnostics
ggplot(emerge,aes(x=ind_type,y=emerg_days,group=assem_type,color=assem_type,fill=assem_type))+geom_violin()
```


```{r}
m1a_emerge <- bglmer(emerg_days~assem_type*ind_type+(1|tank_num),data=emerge,family=Gamma(link="log"))
m1b_emerge <- bglmer(emerg_days~assem_type*ind_type+(1|tank_num),data=emerge,family=gaussian(link="log"))
m1c_emerge <- blmer((emerg_days)~assem_type*ind_type+(1|tank_num),data=emerge)
m1d_emerge<-glmmTMB(emerg_days~assem_type*ind_type+(1|tank_num),data=emerge,family=nbinom1)
library(bbmle)
ICtab(m1a_emerge,m1b_emerge,m1c_emerge,type="AICc")
qqnorm(resid(m1c_emerge))
qqline(resid(m1c_emerge))


```


```{r}
summary(m1c_emerge)
Anova(m1c_emerge, type="III")
```


```{r}
library(broom)
plot(m1c_emerge,idLabels=~.obs,id=.1,title="log(date)")
qqnorm(resid(m1c_emerge))
qqline(resid(m1c_emerge))
sc1=ggplot(augment(m1c_emerge),
       aes(x=.fitted,y=sqrt(abs(.resid/sd(.resid)))))+geom_point()+    
  geom_smooth(colour="red",alpha=0.3)+ggtitle("log(mass")+geom_hline(yintercept = 1)
sc2=ggplot(augment(m1c_emerge),
       aes(x=.fitted,y=.resid))+geom_point()+geom_hline(yintercept = 0)
  
grid.arrange(sc1,sc2,ncol=2)
##Its not perfect but I think as good as we can get.

```


```{r}
#Regression estimates
emerge$ind_type=as.factor(emerge$ind_type)
emerge$assem_type=as.factor(emerge$assem_type)
emerge$trmt_ID=factor(emerge$trmt_ID, levels=c("NL_C_S", "C_C_S", "NL_C_M", "C_C_M"))
str(emerge)

#a look at Regression estimates, with and without random effects. Nothing significant here
library(dotwhisker)
library(coefplot2)
dotwhisker::dwplot(m1c_emerge,effects="fixed",by_2sd=FALSE)
#rr <- "http://www.math.mcmaster.ca/bolker/R"
#install.packages(c("coefplot2"),type="source",
                # repos=rr)
coeftab(m1c_emerge)[,c("Estimate","2.5%","97.5%")]
coefplot2(m1c_emerge)
predict(m1c_emerge,type="response",se=TRUE)
```

```{r}
#predicted values and confidence intervals
dataset3=emerge
boot_predict <- function(.) {
  predict(., newdata=dataset3, re.form=NULL)
}


library(TMB)
bootSummary <- function(merBoot) {
  return(
    data.frame(fit = apply(merBoot$t, 2, function(x) as.numeric(quantile(x, probs=.5, na.rm=TRUE))),
               lwr = apply(merBoot$t, 2, function(x) as.numeric(quantile(x, probs=.025, na.rm=TRUE))),
               upr = apply(merBoot$t, 2, function(x) as.numeric(quantile(x, probs=.975, na.rm=TRUE)))
    )
  )
}



merboot=bootMer(m1c_emerge, boot_predict, nsim=200, use.u=FALSE, type="parametric")
bootstrap.PIs <- bootSummary(merboot)
emerge.frame=data.frame(dataset3,bootstrap.PIs)
library(dplyr)
emerge.frame %>% 
  group_by(ind_type, assem_type) %>% 
  summarise(fit = mean(fit),lwr=mean(lwr),upr=mean(upr)) -> emerge.frame2




library(ggplot2)
ggplot(aes(x=ind_type, y=fit, ymin=lwr, ymax=upr, color=assem_type), data=emerge.frame2) +
  geom_point(size=4,position = position_dodge(w = 0.75))+
  geom_errorbar(position = position_dodge(w = 0.75)) +
  geom_line(aes(group=assem_type),position = position_dodge(w = 0.75))+
  labs(x="Induction Type", y="Time-to Emergence w/95% PI's") +
  theme_bw() +  theme(legend.position="right") +
  scale_color_brewer(type = "qual", palette = 2)
```

```{r}
#Plotting raw data: 
new_emerge <- read.csv(file = "~/Desktop/Dates_Fem.csv",
                             header = TRUE,
                             sep = ",",
                             stringsAsFactors = FALSE)

library(ggplot2)
ggplot(new_emerge, aes(x=assem_type, y=emerg_days, fill=assem_type))+
  geom_boxplot(size=.75, point=TRUE)+
  geom_jitter(aes(colour=assem_type), alpha=0.5)+
  scale_fill_manual(values = c("darkseagreen4","darkorange2"))+
  stat_summary(fun.y = "mean", geom = "point", shape = 15, size = 2)+
  labs(title="Femoralis Time-to Emergence", x="Assemblage Type", y="Time-to Emergence (days)")+
facet_wrap(~ind_type, labeller = labeller(ind_type = ind_labs))+
  scale_color_brewer(type="qual", palette=2)+
  theme_bw()

```



CODE I MAY NEED LATER:

pkgs_CRAN <- c("lme4","MCMCglmm","blme",
               "pbkrtest","coda","aods3","bbmle","ggplot2",
               "reshape2","plyr","numDeriv","Hmisc",
               "plotMCMC","gridExtra","R2admb","pbkrtest")
install.packages(pkgs_CRAN)


rr <- "http://www.math.mcmaster.ca/bolker/R"
install.packages(c("glmmADMB","coefplot2"),type="source",
                 repos=rr)



install.packages("TMB", type="source")
install.packages("glmmTMB", type="source")

library(rhub)
install.packages("glmmTMB", repos="https://glmmTMB.github.io/glmmTMB/repos", type="binary")
