---
title: "Time-to Emergence_Models_042121"
output: html_notebook
---

TIME-TO EMERGENCE Models as of 042121

```{r}
#load packages and data
#loading packages
library(predictionInterval)
library(coefplot2)
library(merTools)
#library(AICcmodavg)
library(brms)
library(utils)
library(dbplyr)
library(TMB)
library(glmmTMB)
library(stats)
library(finalfit)
library(boot)
library(lme4)
library(MCMCglmm)
library(blme)
library(pbkrtest)
library(coda)
library(aods3)
library(bbmle)
library(ggplot2)
library(reshape2)
library(plyr)
library(numDeriv)
library(Hmisc)
library(plotMCMC)
library(gridExtra)
library(R2admb)
library(tibble)
library(tidyverse)
library(car)
library(MASS)
library(stats4)
library(readr)

new_emerge <- read.csv(file = "~/Desktop/Dates_Fem.csv",
                             header = TRUE,
                             sep = ",",
                             stringsAsFactors = FALSE)


#initial view
ggplot(new_emerge,aes(x=ind_type,y=emerg_days,group=assem_type,color=assem_type,fill=assem_type))+geom_violin()
```

```{r}
#running models
m1a_emerge <- bglmer(emerg_days~assem_type*ind_type+(1|tank_num),data=new_emerge,family=Gamma(link="log"))
summary(m1a_emerge)
Anova(m1a_emerge, type="III")
```


```{r}
m1b_emerge <- bglmer(emerg_days~assem_type*ind_type+(1|tank_num),data=new_emerge,family=gaussian(link="log"))
summary(m1b_emerge)
Anova(m1b_emerge, type="III")
```


```{r}
m1c_emerge <- blmer(log(emerg_days)~assem_type*ind_type+(1|tank_num),data=new_emerge)
summary(m1c_emerge)
Anova(m1c_emerge, type="III")

```


```{r}
#CI's and estimates
Confint(m1c_emerge, estimate = TRUE)
```


```{r}
m1d_emerge<-glmmTMB(emerg_days~assem_type*ind_type+(1|tank_num),data=new_emerge,family=nbinom1)
summary(m1d_emerge)
Anova(m1d_emerge, type="III")
```


```{r}
#running blmer without logging the "days" 
m1e_emerge <- blmer(emerg_days~assem_type*ind_type+(1|tank_num),data=new_emerge)
summary(m1e_emerge)
Anova(m1e_emerge, type="III")
```

```{r}
ICtab(m1a_emerge,m1b_emerge,m1c_emerge,m1e_emerge,type="AICc")
```

```{r}
#diagnostics of m1c_emerge (blmer with log(emerg_days))
qqnorm(resid(m1c_emerge))
qqline(resid(m1c_emerge))

library(broom)
sc1=ggplot(augment(m1c_emerge),
       aes(x=.fitted,y=sqrt(abs(.resid/sd(.resid)))))+geom_point()+
    geom_smooth(colour="red",alpha=0.3)+ggtitle("log(mass")+geom_hline(yintercept = 1)
sc2=ggplot(augment(m1c_emerge),
       aes(x=.fitted,y=.resid))+geom_point()+geom_hline(yintercept = 0)
  
grid.arrange(sc1,sc2,ncol=2)
```

```{r}
#diagnostics of m1c_emerge
qqnorm(resid(m1c_emerge))
qqline(resid(m1c_emerge))

```

```{r}
#Predicted values & CI's for m1c_emerge (blmer with log(emerg_days))
dataset5=new_emerge
boot_predict <- function(.) {
  predict(., newdata=dataset5, re.form=NULL)
}


library(TMB)
bootSummary <- function(merBoot) {
  return(
    data.frame(fit = apply(merBoot$t, 2, function(x) as.numeric(quantile(x, probs=.5, na.rm=TRUE))),
               lwr = apply(merBoot$t, 2, function(x) as.numeric(quantile(x, probs=.025, na.rm=TRUE))),
               upr = apply(merBoot$t, 2, function(x) as.numeric(quantile(x, probs=.975, na.rm=TRUE)))
    )
  )
}



merboot=bootMer(m1c_emerge, boot_predict, nsim=200, use.u=FALSE, type="parametric")
bootstrap.PIs <- bootSummary(merboot)
newemerge.frame=data.frame(dataset5,bootstrap.PIs)

library(dplyr)
library(tidyverse)
newemerge.frame %>% 
  group_by(ind_type,assem_type) %>% 
  summarise(fit = mean(fit),lwr=mean(lwr),upr=mean(upr)) -> newemerge.frame2
newemerge.frame2



#renaming factors 
newemerge.frame2$assem_type<-as.factor(newemerge.frame2$assem_type)
levels(newemerge.frame2$assem_type)<-list(Mixed="M",Single="S")
newemerge.frame2$ind_type<-as.factor(newemerge.frame2$ind_type)
levels(newemerge.frame2$ind_type)<-list(Nonlethal="NL", Control="C")
str(newemerge.frame2)

#transforming logged days
newemerge.frame2$expfit <- exp(newemerge.frame2$fit)
newemerge.frame2$explwr <- exp(newemerge.frame2$lwr)
newemerge.frame2$expupr <- exp(newemerge.frame2$upr)
newemerge.frame2



```


```{r}
#plot for m1c_emerge
ggplot(aes(x=ind_type, y=expfit, ymin=explwr, ymax=expupr, color=assem_type), data=newemerge.frame2) +
  geom_point(size=4,position = position_dodge(w = 0.75))+
  geom_errorbar(position = position_dodge(w = 0.75)) +
  geom_line(aes(group=assem_type),position = position_dodge(w = 0.75))+
  labs(colour="Assemblage Type",x="Induction Treatment", y="Time-to Emergence (days) w/ 95% PIs") +
  theme_bw() +  theme(legend.position="right") +
  scale_color_brewer(type = "qual", palette = 2)
  

```



#Plot for m1c_emerge 
library(ggplot2)
ggplot(aes(x=ind_type, y=fit, ymin=lwr, ymax=upr, color=assem_type), data=newemerge.frame2) +
  geom_point(size=4,position = position_dodge(w = 0.75))+
  geom_errorbar(position = position_dodge(w = 0.75)) +
  geom_line(aes(group=assem_type),position = position_dodge(w = 0.75))+
  labs(x="Induction Type", y="Response w/95% PI's") +
  theme_bw() +  theme(legend.position="right") +
  scale_color_brewer(type = "qual", palette = 2)












