---
title: "UPDATE_Morphometrics 042821"
output: html_notebook
---

MORPHOMETRIC ANALYSIS: 92 tadpoles total 
-49 NL tads
-42 C tads

Reference workflow: Emma Sherratt file:///Users/ellie/Downloads/quick_guide_to_geomorph-_introduction.html


Link for referencing "procD.lm" function: https://www.rdocumentation.org/packages/geomorph/versions/3.3.2/topics/procD.lm

Regressions : https://raw.githubusercontent.com/geomorphR/geomorph/Stable/vignettes/geomorph.assistance.Rmd


Geomorph packages as of 04/2021: https://cran.r-project.org/web/packages/geomorph/vignettes/geomorph.functions.html


GMM tutorial: https://rdrr.io/github/katrinajones/qpal/f/vignettes/GMM_Tutorial.Rmd




```{r}
#loading packages
library(geomorph)
library(ggplot2)
library(shapes)
library(svd)
library(scatterplot3d)
library(rgl)
library(MASS)
library(ape)
library(vegan)
library(RRPP)
library(rgl)
library(permute)
library(lattice)

#loading the new text file with discrete variables
femlands <- read.table(file="~/Desktop/New Lateral Tads_Fem.txt",header=TRUE,
stringsAsFactors = FALSE) 

is.numeric(femlands)
```


```{r}
#forcing landmark coordinates to be numeric with as.matrix (minus the first 4 columns)
coords <- as.matrix(femlands[,-(1:4)])
is.numeric(coords)
```

```{r}
#converting discrete variables to factors 
femlands$treatment=as.factor(femlands$treatment)
femlands$tad_ID=as.factor(femlands$tad_ID)
femlands$block_ID=as.factor(femlands$block_ID)
str(femlands)
```

```{r}
#convert our coords matrix into an array (each row contains all coordinates for a single specimen: 21 landmarks per tad)
array_fem <- arrayspecs(coords,21,2)

#just viewing tad_num 1's coordinate array to see if it worked: 
array_fem[,,1]

```

```{r}
#perform a GPA
library(geomorph)
gpa1 <- gpagen(array_fem, 
       curves = NULL, 
       surfaces = NULL,
       PrinAxes = TRUE,
       max.iter = NULL,
       ProcD = FALSE,
       Proj = TRUE,
       print.progress = TRUE
       )
print(gpa1)
```

```{r}
#viewing the Procrustes analysis results on a grid
plotAllSpecimens(array_fem)
plot(gpa1)

```

```{r}
#create a new dataframe with GPA results & discrete variables
gdf <- geomorph.data.frame(gpa1, treatment=femlands$treatment, tank_num=femlands$tad_ID)
gdf
```

```{r}
#Initial model + ANOVA 

#mod1: Procrustes analysis coordinates & treatment

#RRPP = Randomized Residual permutation procedure: Should the argument "RRPP=" be TRUE or FALSE?
mod1 <- procD.lm(coords~treatment, data=gdf, iter=999, RRPP=TRUE, print.progress=FALSE)
anova(mod1)

```


```{r}
#running a PCA: gives us a plot and principle components, eigenvalues
PCA1 <- gm.prcomp(gpa1$coords)
summary(PCA1)

#plotting the PCA
plot(PCA1, main="PCA1")

#plotting PCA scores for each specimen
#outlier near .20, color the points
plot(PCA1$x)

```

```{r}
#looking at outliers of PCA plot, looks like it's specimen #30 that's the major outlier?
out_PCA1 <- plot(mod1, type="diagnostics", outliers=TRUE)
out_PCA1

```


```{r}
#removing specimen #30 from the array and re-running GPA, PCA plots
array_fem_new <- array_fem[,,-30]
array_fem_new

#re-running the GPA
library(geomorph)
gpa_new <- gpagen(array_fem_new, 
       curves = NULL, 
       surfaces = NULL,
       PrinAxes = TRUE,
       max.iter = NULL,
       ProcD = FALSE,
       Proj = TRUE,
       print.progress = TRUE
       )
print(gpa_new)

#plotting new Procrustes coordinates
plotAllSpecimens(array_fem_new)
plot(gpa_new)

#checking outliers to make sure specimen #30 does not pop up/was indeed removed
plotOutliers(gpa_new$coords, inspect.outliers = TRUE)

#new geomorph data frame without specimen 30
femlands_new <- femlands[-30,]
femlands_new

gdf_new <- geomorph.data.frame(gpa_new, treatment=femlands_new$treatment, tank_num=femlands_new$tad_ID)
gdf_new
```


```{r}
#re-running model, so we can use in new PCA plots
#PERMUTATIONAL Anova that I report: 
mod1_new <- procD.lm(coords~treatment, data=gdf_new, iter=999, RRPP=TRUE, print.progress=FALSE)
summary(mod1_new)
anova.lm.rrpp(mod1_new)
```

```{r}
#new PCA plot without specimen #30
PCA_new <- gm.prcomp(gpa_new$coords)
summary(PCA_new)

#plotting the PCA
plot(PCA_new, main="PCA Without Specimen 30")

#plotting PCA scores for each specimen
plot(PCA_new$x)
```


Code to run MANOVA's on PC scores: 

-Investigating the affect of 1 independent variable (treatment) on multiple dependent variables (multiple PC scores)

http://rstudio-pubs-static.s3.amazonaws.com/8311_eef838dfa3d14f84ba1bab795292990f.html


```{r}
#running a MANOVA : affect of treatment on PC scores 1-5 containing 66.47% of the variance

#Using base R PCA function
gpa_new_2d <- two.d.array(gpa_new$coords)
gpa_new_wholePCA <- prcomp(gpa_new_2d)

#Organizing PCA component scores into a dataframe with treatment type
gpa_newPCA <- data.frame(gpa_new_wholePCA$x[,1],gpa_new_wholePCA$x[,2],gpa_new_wholePCA$x[,3],gpa_new_wholePCA$x[,4],gpa_new_wholePCA$x[,5])
gpa_newPCA <- data.frame(gpa_newPCA,femlands_new$treatment)
names(gpa_newPCA) <- c("PC1","PC2","PC3","PC4","PC5","treatment")


```

```{r}
#trying permutational anova's....not sure if this is what I want?

#Permutational Multivariate Analysis of Variance using distance matrices.....
#----not what I'm doing below.....too simple of a procedure
#need to calculate procrustes distances with coordinates...didn't do so here
#can choose from a multitude of methods

permanov <- adonis(formula=cbind(PC1,PC2,PC3,PC4,PC5)~treatment, data=gpa_newPCA, method="euclidean", permutations = 999)
print(permanov)
```


```{r}
#trying another permutational anova

#creating a data frame with a geometric morphometric measure of shape difference (Centroid sizes) & our predictor (treatment)
newdf <- as.data.frame(gdf_new$Csize)
newdf <- data.frame(cbind(newdf, gdf_new$treatment))
newdf
names(newdf) <- c("Csize","treatment")


permanov2 <- adonis(Csize~treatment, data=newdf, method="euclidean", permutations=999)
print(permanov2)
```


```{r}
#translating PC scores back into original shape variables (landmarks)
pcscores <- PCA_new$x
eigenvectors <- PCA_new$rotation
#concensus shape
ref_main <- mshape(gdf_new$coords)

#calculating concensus shape 
scores <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0) #pc scores of concensus shape
shape <- scores*eigenvectors
shape <- matrix(shape,nrow=21,ncol=2,byrow=T) #will all be zeroes for concensus shape
shape

shape_new <- shape+ref_main
shape_new

```


```{r}
#Now, can calculate the shape at any point in morphospace. Example = What does a positive loading of 0.5 on PC1 look like?
#trying to  investigate what landmarks/aspects of morphology relate to which components
#can substitute values in for any of the 38 components and look at the plot to see what it does to morphology 
scores_2 <- c(0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
shape <- scores_2*eigenvectors
shape <- matrix(shape,nrow=21,ncol=2,byrow=T)

shape <- shape+ref_main
plotRefToTarget(ref_main,shape)
```


```{r}
#Creating subsets of landmark data for each treatment type: Data-wrangling
data(femlands_new)
colnames(femlands_new)

library(tidyverse)
library(dplyr)
group_NL <- femlands_new %>%
  filter(treatment=="NL") %>%
  select(tad_ID, treatment, block_ID, tad_num, X1X:X21Y)

group_NL

group_C <- femlands_new %>%
        filter(treatment=="C") %>%
        select(tad_ID, treatment, block_ID, tad_num, X1X:X21Y)

group_C

#arrays
coords_NL <- as.matrix(group_NL[,-(1:4)])
is.numeric(coords_NL)
coords_C <- as.matrix(group_C[,-(1:4)])
is.numeric(coords_C)


array_femNL <- arrayspecs(coords_NL,21,2)
array_femNL[,,1]

array_femC <- arrayspecs(coords_C,21,2)
array_femC[,,1]
```

```{r}
#running GPA's on each treatment type: creating shape variables
gpa2 <- gpagen(array_femNL, 
       curves = NULL, 
       surfaces = NULL,
       PrinAxes = TRUE,
       max.iter = NULL,
       ProcD = FALSE,
       Proj = TRUE,
       print.progress = TRUE
       )
print(gpa2)

gpa3 <- gpagen(array_femC, 
       curves = NULL, 
       surfaces = NULL,
       PrinAxes = TRUE,
       max.iter = NULL,
       ProcD = FALSE,
       Proj = TRUE,
       print.progress = TRUE
       )
print(gpa3)


```

```{r}
library(graphics)
#NL GPA
plot(gpa2)
#C GPA
plot(gpa3)
```



```{r}
#calculating mean shape from each GPA for each treatment group
NL <- gpa2$coords
ref1 <- mshape(NL, na.action=1)
ref1

C <- gpa3$coords
ref2 <- mshape(C, na.action=1)
ref2
```

```{r}
#concensus shape = mshape
ref_main <- mshape(gpa_new$coords)
plotRefToTarget(ref_main,gpa_new$coords[,,91])


#Looking at treatment group means vs. treatment group Procrustes coordinates for all treatment specimens

#I think this shows that NL tads overall had smaller heads, deeper tail fins (distance from dorsal to ventral edge), and perhaps slightly longer tails?
ref_NL <- mshape(gpa2$coords)
plotRefToTarget(ref_NL,gpa2$coords[,,49])

#C tads seem to have an overall proportionate body shape: heads are larger in comparison to NL, and tails not as deep in relation to head size
ref_C <- mshape(gpa3$coords)
plotRefToTarget(ref_C,gpa3$coords[,,42])

```



https://github.com/cran/geomorph/blob/master/R/plotRefToTarget.r



```{r}
#You can play around with the "mag" & "method" parameter

#first term = reference specimen, second term=target specimen
#ref1 = NL, ref2 = C


#playing with colors of points - these might be nice overlay images for the PCA plots
#reference specimen is highlighted in red, target is black 
plotRefToTarget(ref1, ref2, method="points",mag=5,gridPars=gridPar(pt.bg = "red", pt.size = 1.25))

#Making ref2 (Control tads) the reference specimen, and ref1 (NL tads) the target
plotRefToTarget(ref2, ref1, method="points",mag=4,gridPars=gridPar(pt.bg = "lightgrey", pt.size = 1.25))

#vector method
#reference specimen (gray dots with arrows) moving towards target specimen shape (black dots)

#overall concensus shape, moving towards mean NL treatment group shape
plotRefToTarget(ref_main,ref_NL,method=c("vector"),mag=5)
plotRefToTarget(ref_main,ref_NL,method=c("vector"),mag=10)
plotRefToTarget(ref_main,ref_NL,method=c("vector"),mag=20)


#overall concensus shape, moving towards mean NL treatment group shape
plotRefToTarget(ref_main,ref_C,method=c("vector"),mag=20)

```
Grey = NL tads
Black =  C tads


Reference this site: 

http://creativemorphometrics.co.vu/blog/2014/02/23/r-generator-and-a-colorful-pca/



```{r}
#plotting PCA's of each treatment type just to see
PCA_NL <- gm.prcomp(gpa2$coords)
PCA_NL
plot(PCA_NL, main="NL")
PCA_C <- gm.prcomp(gpa3$coords)
PCA_C
plot(PCA_C, main="C")
```


```{r}
#Calculating mean PCA scores for PC1-PC5 per treatment type
meangpa_newPCA1 <- aggregate(gpa_newPCA[,1], mean, by=list(gpa_newPCA[,6]))
names(meangpa_newPCA1)<-c("treatment","PC1_score")
meangpa_newPCA2 <- aggregate(gpa_newPCA[,2], mean, by=list(gpa_newPCA[,6]))
names(meangpa_newPCA2)<-c("treatment","PC2_score")
meangpa_newPCA3 <- aggregate(gpa_newPCA[,3], mean, by=list(gpa_newPCA[,6]))
names(meangpa_newPCA3)<-c("treatment","PC3_score")
meangpa_newPCA4 <- aggregate(gpa_newPCA[,4], mean, by=list(gpa_newPCA[,6]))
names(meangpa_newPCA4)<-c("treatment","PC4_score")
meangpa_newPCA5 <- aggregate(gpa_newPCA[,5], mean, by=list(gpa_newPCA[,6]))
names(meangpa_newPCA5)<-c("treatment","PC5_score")
meangpa_newPCA <- data.frame(meangpa_newPCA1,meangpa_newPCA2[,2],meangpa_newPCA3[,2],meangpa_newPCA4[,2],meangpa_newPCA5[,2])
names(meangpa_newPCA) <- c("treatment","PC1","PC2","PC3","PC4","PC5")


#plotting
#pcaplot <- pcaplot + geom_point(data = meangpa_newPCA,size=5,aes(color=treatment,shape=treatment)) + geom_text(data = meangpa_newPCA, size = 5, label = meangpa_newPCA$treatment, vjust = -0.9) 
#pcaplot

library (ggplot2)
library(ggforce)
data=gpa_newPCA
pcaplot <- ggplot(gpa_newPCA, aes(x=PC1, y=PC2, group = treatment, shape=treatment)) + geom_point(size = 1.25, aes(color=treatment)) +
     scale_color_manual(name="Treatments", 
      labels=c(
        C="Control",
      NL="Non-Lethal"
      ),                  
      values=c(
      C="black",
      NL ="red"))+
  xlab("PC1 : 23.99%")+
  ylab("PC2 : 14.37%")+
  geom_vline(xintercept = 0.0,linetype="dashed",color="grey50")+
  geom_hline(yintercept = 0.0,linetype="dashed",color="grey50")+
  theme_bw()
pcaplot <- pcaplot + geom_point(data = meangpa_newPCA,size=5,aes(color=treatment,shape=treatment)) + geom_text(data = meangpa_newPCA, size = 5, label = meangpa_newPCA$treatment, vjust = -0.9) 
pcaplot

```


```{r}
pcaplot2 <- ggplot(gpa_newPCA, aes(x=PC1, y=PC2, group = treatment, shape=treatment)) + 
  geom_point(size = 1.25, aes(color=treatment))+
       scale_color_manual(values=c(
      C="black",
      NL ="red"))+
  geom_mark_ellipse(expand=0,aes(color=treatment))+
  xlab("PC1 : 23.99%")+
  ylab("PC2 : 14.37%")+
  theme_bw()

pcaplot2 <- pcaplot2 + geom_point(data = meangpa_newPCA,size=5,aes(color=treatment,shape=treatment)) + 
  geom_text(data = meangpa_newPCA, size = 5, label = meangpa_newPCA$treatment, vjust = -0.9) 
pcaplot2

```

```{r}
#more plots
pcaplot3 <- ggplot(gpa_newPCA, aes(x=PC1, y=treatment, group = treatment, shape=treatment)) + 
  geom_point(size = 2, aes(color=treatment))+
       scale_color_manual(values=c(
      C="black",
      NL ="red"))+
  scale_x_continuous(limits=c(-0.09,0.09))+
  theme_bw()
pcaplot3<- pcaplot3 + geom_point(data = meangpa_newPCA,size=5,aes(color=treatment,shape=treatment)) + 
  geom_text(data = meangpa_newPCA, size = 5, label = meangpa_newPCA$treatment, vjust = -0.9)
pcaplot3



pcaplot4 <- ggplot(gpa_newPCA, aes(x=PC2, y=treatment, group = treatment, shape=treatment)) + 
  geom_point(size = 2, aes(color=treatment))+
       scale_color_manual(values=c(
      C="black",
      NL ="red"))+
  scale_x_continuous(limits=c(-0.09,0.09))+
  theme_bw()
pcaplot4 <- pcaplot4 + geom_point(data = meangpa_newPCA,size=5,aes(color=treatment,shape=treatment)) + 
  geom_text(data = meangpa_newPCA, size = 5, label = meangpa_newPCA$treatment, vjust = -0.9)
pcaplot4
```
```{r}
#PC3 & 4 ellipses
pcaplot_34 <- ggplot(gpa_newPCA, aes(x=PC3, y=PC4, group = treatment, shape=treatment)) + 
  geom_point(size = 1.25, aes(color=treatment))+
       scale_color_manual(values=c(
      C="black",
      NL ="red"))+
  geom_mark_ellipse(expand=0,aes(color=treatment))+
  xlab("PC3 : 12.13%")+
  ylab("PC4 : 8.63%")+
  theme_bw()

pcaplot_34 <- pcaplot_34 + geom_point(data = meangpa_newPCA,size=5,aes(color=treatment,shape=treatment)) + geom_text(data = meangpa_newPCA, size = 5, label = meangpa_newPCA$treatment, vjust = -0.9) 
pcaplot_34
```


```{r}
#plotting PC4 and 5
pcaplot_45 <- ggplot(gpa_newPCA, aes(x=PC4, y=PC5, group = treatment, shape=treatment)) + 
  geom_point(size = 1.25, aes(color=treatment))+
       scale_color_manual(values=c(
      C="black",
      NL ="red"))+
  geom_mark_ellipse(expand=0,aes(color=treatment))+
  xlab("PC4 : 8.63%")+
  ylab("PC5 : 7.35%")+
  theme_bw()

pcaplot_45 <- pcaplot_45 + geom_point(data = meangpa_newPCA,size=5,aes(color=treatment,shape=treatment)) + geom_text(data = meangpa_newPCA, size = 5, label = meangpa_newPCA$treatment, vjust = -0.9) 
pcaplot_45
```


```{r}
#Sys.setenv(R_REMOTES_NO_ERRORS_FROM_WARNINGS="true")
#library(devtools)
#install_github("vqv/ggbiplot")

#library(ggbiplot)



#ggbiplot(gpa_newPCA,ellipse = TRUE,choices=c(1,2),labels=colnames(gpa_newPCA),groups=gpa_newPCA$treatment)

```

