---
title: "UPDATED_Survival Analysis"
output: html_notebook
---

UPDATED SURVIVAL ANALYSIS as of 5/5/21



Consolidated all survival data into one master csv

-Status at time t = 1 = Alive, 0 = Dead



First, loading the data

```{r}
survival <- read.csv(file = "~/Desktop/Survival_MASTER.csv",
                             header = TRUE,
                             sep = ",",
                             stringsAsFactors = FALSE)

library(ggplot2)
library(tidyverse)
library(dplyr)
library(knitr)
library(lubridate)
library(survival)
library(survminer)
```

Adding a column for proportion surviving at time points t: 

```{r}
survival <- survival %>%
  mutate(
    prop_surv = tads_remain/tads_start
  )
```

Next, creating subset-data frames and plotting each interesting comparison with "Proportion Surviving" on the Y axis: 

```{r}
single_NLvC <- survival[1:160, 1:13]

library(ggplot2)
colors <- c("C_L_S" = "black", "NL_L_S" = "red")
ggplot(single_NLvC, aes(x = time, y = prop_surv, color = ind_ID)) +
    geom_point(size = .25) + 
    scale_color_manual(values = colors, labels = c("C femoralis", "NL femoralis")) +
    stat_smooth(method = "glm", method.args = list(family = "quasibinomial"), se = FALSE) +
    labs(x = "Time (Hours)", y = "Proportion Surviving", title = "Femoralis Survival = Single-Sp. Assemblages") + 
    theme_bw()
```

```{r}
mixed_NLvC <- survival[161:320, 1:13]

library(ggplot2)
colors <- c("C_L_M" = "black", "NL_L_M" = "red")
ggplot(mixed_NLvC, aes(x = time, y = prop_surv, color = ind_ID)) +
    geom_point(size = .25) + 
    scale_color_manual(values = colors, labels = c("C femoralis", "NL femoralis")) +
    stat_smooth(method = "glm", method.args = list(family = "quasibinomial"), se = FALSE) +
    labs(x = "Time (Hours)", y = "Proportion Surviving", title = "Femoralis Survival = Mixed-Sp. Assemblages (4 TADS)") + 
    theme_bw()
```



RUNNING Initial model on FEMORALIS SURVIVAL:


First, create a dataframe that contains only femoralis survival (across assemblages & induction types) :


Now, for my GLM I have: 

prop-surv = bound by 0 & 1 = continuous

fixed effects = time, ind_type (NL or C) and assem_type (S or M)
random effects = tank_num and (possibly?) block_ID

error family = binomial (alive or dead)

- I have column "status" (says whether you're 1 = alive, 0 = dead)...
-Going to have "time" as a fixed effect, as the time point (hours passed) relates to your status (alive or dead)


Mike edits from 3/4/21: 

```{r}
names(survival)

survdat <- survival %>%
  filter(
    species =="Fe"
  )

#modeling survival
library(broom)
library(AICcmodavg)
library(lme4)
library(car)
library(nnet)
library(MASS)


#OUR MODEL: with random effect of block_ID to make boot code work below
library(blme)
m2<-glmer(prop_surv~ind_ID*time+(1|block_ID),family=binomial,weights=tads_start,data=survdat)
summary(m2)
Anova(m2,type="III")
```


```{r}
cis <- Confint(m2, estimate = TRUE)
cis
```

Converting logit (coefficient) values to plogis scale (y probabilities):

https://sebastiansauer.github.io/convert_logit2prob/



```{r}
#converting coefficients
plogis(0.006)

```


```{r}
#plotting RAW data
library(ggplot2)       
ggplot(data=survdat,aes(x=time,y=prop_surv,color=ind_ID, linetype=ind_ID))+
    scale_color_manual(name="Treatments", 
      labels=c("Control Mixed",
               "Control Single",
               "Non-lethal Mixed",
               "Non-lethal Single"),
      values=c(
      C_L_S="black",
      C_L_M="azure4",
      NL_L_S="darkred",
      NL_L_M="brown1"))+
    scale_linetype_manual(values=c(
      C_L_S="solid",
      C_L_M="longdash",
      NL_L_S="solid",
      NL_L_M="longdash"))+
    geom_point(size=0.25)+geom_smooth(size=1.15, method = "glm", 
    method.args = list(family = "binomial"), 
    se = FALSE)+
    labs(title="Femoralis Survival")+
    xlab("Time (Hours")+
    ylab("Proportion Surviving")+
    theme_light()
```
    
```{r}
#predicted values from our model m2
library(dplyr)

survdat$ind_ID<-as.factor(survdat$ind_ID)
survdat$tank_num<-as.factor(survdat$tank_num)
survdat$block_ID<-as.factor(survdat$block_ID)




dataset=survdat
boot_predict <- function(.) {
  predict(., newdata=dataset, re.form=NULL)
}


#install.packages('TMB', type = 'source')
bootSummary <- function(merBoot) {
  return(
    data.frame(fit = apply(merBoot$t, 2, function(x) as.numeric(quantile(x, probs=.5, na.rm=TRUE))),
               lwr = apply(merBoot$t, 2, function(x) as.numeric(quantile(x, probs=.025, na.rm=TRUE))),
               upr = apply(merBoot$t, 2, function(x) as.numeric(quantile(x, probs=.975, na.rm=TRUE)))
    )
  )
}

merboot=bootMer(m2, boot_predict, nsim=1000, use.u=FALSE, type="parametric")
bootstrap.PIs <- bootSummary(merboot)
survframe=data.frame(dataset,bootstrap.PIs)


library(dplyr)
survframe %>% 
  group_by(ind_ID,time) %>% 
  summarise(fit = mean(fit),lwr=mean(lwr),upr=mean(upr)) -> survframe2
survframe2

```


```{r}
#adding column of assemblage type to model output
library(dplyr)
assemblage<-c("Mixed","Single")
assemblages<-data.frame(assemblage)

assemblages_new<-data.frame(assemblages[rep(seq_len(nrow(assemblages)),each=16),])
assemblages_new
names(assemblages_new)<-c("assemblage")

assemblages_new2<-data.frame(assemblages[rep(seq_len(nrow(assemblages)),each=16),])
assemblages_new2
names(assemblages_new2)<-c("assemblage")

assem_TOTAL <- data.frame(cbind(assemblages_new$assemblage,assemblages_new2$assemblage))
assem_TOTAL

assem_long <- data.frame(gather(assem_TOTAL,assemblage))
assem_long
names(assem_long)<-c("value","assem_type")

final_survframe<-data.frame(cbind(survframe2,assem_long$assem_type))
final_survframe
names(final_survframe)<-c("ind_ID","time","fit","lwr","upr","assem_type")

```


```{r}
#PLOTTING MODEL OUTPUT
library(ggplot2)
ggplot(aes(x=time,y=plogis(fit),ymin=plogis(lwr),ymax=plogis(upr),color=ind_ID,linetype=ind_ID),data=final_survframe)+
  stat_summary(fun="mean",size=1,geom="line")+
     scale_color_manual(name="Treatments", 
      labels=c("Control Mixed",
               "Control Single",
               "Non-lethal Mixed",
               "Non-lethal Single"),
      values=c(
      C_L_S="black",
      C_L_M="azure4",
      NL_L_S="darkred",
      NL_L_M="brown1"))+
    scale_linetype_manual(values=c(
      C_L_S="solid",
      C_L_M="longdash",
      NL_L_S="solid",
      NL_L_M="longdash"))+
  labs(x="Time (hrs)",y="Proportion Surviving w/95% PI's")+
  theme_bw()+theme(legend.position="right")
```








```{r}
#intercepts & slope
coef(m2)
```


```{r}
#wrangling final survframe
levels(final_survframe$ind_ID)<-list(Control_Mixed = "C_L_M", Control_Single = "C_L_S", Nonlethal_Mixed = "NL_L_M", Nonlethal_Single = "NL_L_S")
names(final_survframe)
```


```{r}
#MORE MODEL PLOTS : FACETS 
library(ggplot2)
ggplot(aes(x=time,y=plogis(fit),ymin=plogis(lwr),ymax=plogis(upr),color=ind_ID, linetype=ind_ID),data=final_survframe)+
  geom_ribbon(aes(alpha=.1),fill="grey80")+geom_line(size=1)+
     scale_color_manual(
       values=c(
      Control_Single="black",
      Control_Mixed="azure4",
      Nonlethal_Single="darkred",
      Nonlethal_Mixed="brown1"))+
    scale_linetype_manual(values=c(
      Control_Single="solid",
      Control_Mixed="dashed",
      Nonlethal_Single="solid",
      Nonlethal_Mixed="dashed"))+
  labs(x="Time (hrs)",y="Proportion Surviving w/95% PI's")+
  theme_bw()+theme(legend.position="right")+
  facet_wrap(~ind_ID)
```


```{r}
```{r}